NTM CT Scoring Agreement

### 2025-08-08 Stacey Martiniano/ntm_agreement.qmd
need to ask Stacey about the total score thing
NTM added score totals and scatterplots

```{r}
#| include: false
library(tidyverse)
library(redcapAPI)
library(Hmisc)
library(knitr)
library(lmerTest)
library(performance)
# REDCap
unlockREDCap(c(rcon = "NTM CT Scoring"),
  keyring = "API_KEYs",
  envir = globalenv(),
  url = "https://redcap.ucdenver.edu/api/"
)
# Home directory
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

## Total disease feature score (orange)

```{r}
# Import disease state
outcomes <-
  read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",
    na.strings = ""
  )
outcomes$NTM.PD..Y.N. <- factor(outcomes$NTM.PD..Y.N.,
  levels = c("No", "Yes"), ordered = T
)
outcomes <- outcomes %>%
  select(cff_registry_id, NTM.PD..Y.N.) %>%
  rename(`NTM-PD` = NTM.PD..Y.N.)
# For the duplicated scans, just use the first reading and combine with the
# singletons
scores <- rbind(data, data_dups[data_dups$scorer_repeat == 1, ])
# Add gold standards
scores <- rbind(scores, data_gs[data_gs$scorer == "GS", ])
# Calculate the scores for each feature.
# When calculating the scores, take the majority vote by scan, feature, and
# location, since that's how they scored it.
# For example: If one scorer detected atelectasis in the RUL, one in the RML,
# and one in the RLL, but the majority at each location was "absent," then that
# scan would have a 0 for atelectasis. In other words, score the locations by
# majority first, then add the locations together for the feature score.
feature_scores <- scores %>%
  group_by(scan_id, Feature, Location) %>%
  summarise(Majority = factor(sum(value == "Present"),
    levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
  ), .groups = "drop_last") %>%
  summarise(Score = sum(Majority == "Present"), .groups = "drop")
# Add outcome data
feature_scores <- left_join(feature_scores, outcomes,
  by = join_by(scan_id == cff_registry_id)
)
# Summary table
feature_scores %>%
  pivot_wider(names_from = Feature, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(by = `NTM-PD`, missing_text = "Missing")
feature_scores %>%
  pivot_wider(names_from = Feature, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(
    by = `NTM-PD`, missing_text = "Missing",
    type = list(
      c(
        "Atelectasis", "Bronchiectasis", , "Cavity", "Consolidation", "Cyst",
        "Ground Glass", "Mucoid Impaction", "Nodule", "Tree In Bud",
        "Tree In Bud OR Nodule"
      ) ~ "continuous2"
    ),
    statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
  )
```

>>>>

```{r}
#| include: false
library(tidyverse)
library(redcapAPI)
library(Hmisc)
library(knitr)
library(lmerTest)
library(performance)
library(gtsummary)
library(patchwork)
# REDCap
unlockREDCap(c(rcon = "NTM CT Scoring"),
  keyring = "API_KEYs",
  envir = globalenv(),
  url = "https://redcap.ucdenver.edu/api/"
)
# Home directory
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF",
  "Windows" = "C:/Users/Tim/OneDrive - The University of Colorado Denver/Vigers/CF",
  "Linux" = "/home/tim/OneDrive/Vigers/CF"
)
github_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/tim/Documents/GitHub",
  "Windows" = "C:/Users/Tim/Documents/GitHub",
  "Linux" = "/home/tim/Documents/GitHub"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

```{r}
# Import disease state
outcomes <-
  read.csv("./Stacey Martiniano/NTM CT Scoring/Data_Clean/disease.key.csv",
    na.strings = ""
  )
outcomes$NTM.PD..Y.N. <- factor(outcomes$NTM.PD..Y.N.,
  levels = c("No", "Yes"), ordered = T
)
outcomes <- outcomes %>%
  select(cff_registry_id, NTM.PD..Y.N.) %>%
  rename(`NTM-PD` = NTM.PD..Y.N.)
# For the duplicated scans, just use the first reading and combine with the
# singletons
scores <- rbind(data, data_dups[data_dups$scorer_repeat == 1, ])
# Add gold standards
scores <- rbind(scores, data_gs[data_gs$scorer == "GS", ])
```

## Feature score (orange)

```{r}
# Calculate the scores for each feature.
# When calculating the scores, take the majority vote by scan, feature, and
# location, since that's how they scored it.
# For example: If one scorer detected atelectasis in the RUL, one in the RML,
# and one in the RLL, but the majority at each location was "absent," then that
# scan would have a 0 for atelectasis. In other words, score the locations by
# majority first, then add the locations together for the feature score.
feature_scores <- scores %>%
  group_by(scan_id, Feature, Location) %>%
  summarise(Majority = factor(sum(value == "Present", na.rm = T),
    levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
  ), .groups = "drop_last") %>%
  summarise(Score = sum(Majority == "Present"), .groups = "drop")
# Add outcome data
feature_scores <- left_join(feature_scores, outcomes,
  by = join_by(scan_id == cff_registry_id)
)
# Summary table
feature_scores %>%
  pivot_wider(names_from = Feature, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(by = `NTM-PD`, missing_text = "Missing")
feature_scores %>%
  pivot_wider(names_from = Feature, values_from = Score) %>%
  mutate(`Total Score` = rowSums(.[3:12])) %>%
  select(-scan_id) %>%
  tbl_summary(
    by = `NTM-PD`, missing_text = "Missing",
    type = list(
      c(
        "Atelectasis", "Bronchiectasis", , "Cavity", "Consolidation", "Cyst",
        "Ground Glass", "Mucoid Impaction", "Nodule", "Tree In Bud",
        "Tree In Bud OR Nodule", "Total Score"
      ) ~ "continuous2"
    ),
    statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
  ) %>%
  add_p(
    test.args = all_tests("wilcox.test") ~ list(exact = F)
  )
```

## Lobe score (pink)

```{r}
# Lobe scores similar to above but we'll add over the lobes rather than the
# features
lobe_scores <- scores %>%
  group_by(scan_id, Location, Feature) %>%
  summarise(Majority = factor(sum(value == "Present", na.rm = T),
    levels = 0:3, labels = c("Absent", "Absent", "Present", "Present")
  ), .groups = "drop_last") %>%
  summarise(Score = sum(Majority == "Present"), .groups = "drop")
# Add outcome data
lobe_scores <- left_join(lobe_scores, outcomes,
  by = join_by(scan_id == cff_registry_id)
)
# Summary table
lobe_scores %>%
  pivot_wider(names_from = Location, values_from = Score) %>%
  select(-scan_id) %>%
  tbl_summary(by = `NTM-PD`, missing_text = "Missing")
lobe_scores %>%
  pivot_wider(names_from = Location, values_from = Score) %>%
  mutate(`Total Score` = rowSums(.[3:8])) %>%
  select(-scan_id) %>%
  tbl_summary(
    by = `NTM-PD`, missing_text = "Missing",
    type = list(
      c("RUL", "RML", "RLL", "LUL", "Ling", "LLL", "Total Score") ~ "continuous2"
    ),
    statistic = all_continuous() ~ c("{median} ({p25}, {p75})", "{min}, {max}")
  ) %>%
  add_p(
    test.args = all_tests("wilcox.test") ~ list(exact = F)
  )
```

# Scatterplots

```{r}
#| warning: false
#| fig-width: 12
#| fig-height: 9
# For the duplicated scans, just use the first reading and combine with the
# singletons
scores <- rbind(data, data_dups[data_dups$scorer_repeat == 1, ])
# Add gold standards
scores <- rbind(scores, data_gs[data_gs$scorer != "GS", ])
# Get a total by scorer
plot_df <- scores %>%
  group_by(scan_id, scorer) %>%
  summarise(sum = sum(value == "Present"), .groups = "drop") %>%
  pivot_wider(names_from = "scorer", values_from = sum)
# Plots
ab <- plot_df %>%
  ggplot(., aes(x = A, y = B)) +
  geom_point() +
  ggtitle("Total Score") +
  geom_abline(intercept = 0, slope = 1) +
  coord_fixed(ratio = 1, xlim = c(0, 45), ylim = c(0, 45)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
ac <- plot_df %>%
  ggplot(., aes(x = A, y = C)) +
  geom_point() +
  ggtitle("Total Score") +
  geom_abline(intercept = 0, slope = 1) +
  coord_fixed(ratio = 1, xlim = c(0, 45), ylim = c(0, 45)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
bc <- plot_df %>%
  ggplot(., aes(x = B, y = C)) +
  geom_point() +
  ggtitle("Total Score") +
  geom_abline(intercept = 0, slope = 1) +
  coord_fixed(ratio = 1, xlim = c(0, 45), ylim = c(0, 45)) +
  theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))
ab + bc + ac
```
